---
title: "Illusion Sensitivity and Phenomenological Control"
subtitle: "Analysis - Departures From Preregistration"
date: "`r Sys.Date()`"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: true
    toc: true
    warning: false 
    message: false 
execute: 
  cache: false
---

## Data Preparation 

```{r}
library(tidyverse)
library(ggdist)
library(ggside)
library(easystats)
library(patchwork)
library(psych)


data_ppt <- read.csv("C:/Users/olive/Documents/University/Year 3 (Placement Year)/REBEL/IllusionGamePhenomenologicalControl/Data/data_participants.csv")

data_ig <- read.csv("C:/Users/olive/Documents/University/Year 3 (Placement Year)/REBEL/IllusionGamePhenomenologicalControl/Data/data_illusion.csv") |>
    mutate(
        Block = as.factor(Block),
        Illusion_Effect = fct_relevel(as.factor(Illusion_Effect), "Incongruent", "Congruent"),
        Response_Correct = as.factor(Response_Correct)
    )

```

## PCS Exclusion

```{r}
outliers_pcs1 <- data_ppt[data_ppt$pcs_hello != "Hello", "Participant"]
outliers_pcs2 <- data_ppt[data_ppt$pcs_noballs == TRUE, "Participant"]
outliers_pcs3 <- data_ppt[data_ppt$pcs_press < 5, "Participant"] # get PP count
outliers_pcs <- unique(c(outliers_pcs1, outliers_pcs2))
```

Departing from our preregistered analysis, here we will not include the third attention check, in which participants were told to press the space bar upon seeing the words "PLEASE WAIT" on the screen. This is because a significant proportion (`r length(outliers_pcs3)` (`r insight::format_percent(length(outliers_pcs3) / nrow(data_ppt))`)) of participants were excluded. Thus, we will only remove `r length(outliers_pcs)` (`r insight::format_percent(length(outliers_pcs) / nrow(data_ppt))`) participants that either did not report hearing "Hello" at the beginning of the experiment, or responded that no balls were presented on the screen.

```{r}
data_ppt <- filter(data_ppt, !Participant %in% outliers_pcs)
data_ig <- filter(data_ig, !Participant %in% outliers_pcs)
```

::: {.panel-tabset}

### PCS Exclusions Table

```{r}
pcs_ex_tib <- bind_rows(
  tibble(Participant = outliers_pcs1, Reason = "Failed 'Hello' check"),
  tibble(Participant = outliers_pcs2, Reason = "Reported no balls"),
)

pcs_ex_tib |> 
  knitr::kable(caption = "Exclusions From Phenomenological Control Scale") |> 
  kableExtra::scroll_box(height = "300px") |> 
  kableExtra::kable_styling()
```

### PCS Exclusions Graph

```{r}
pcs_ex_bar <- pcs_ex_tib |> 
  ggplot(aes(x = Reason, fill = Reason)) +
  geom_bar() +
  scale_y_continuous(breaks = scales::pretty_breaks()) +
    labs(
    title = "Exclusions From Phenomenological Control Scale",
    x = "Reason for Exclusion",
    y = "Count",
  ) +
  theme_minimal() +
  guides(fill = "none")

pcs_ex_bar
```

:::

## Score Computation

### PCS

Mean subjective scale responses for item 4 (Taste) and item 10 (Post-session suggestion) are each generated from the combination of two responses which require processing prior to generating an overall subjective score. 
The subjective score for item 4 is the mean of the responses to the (a)‘sour’ and (b)‘sweet’ sub-scales for the taste suggestion. 
The subjective score for item 10 is the geometric mean (the square root of the product) of the responses to the (a) ‘urge’ and (b) ‘amnesia’ sub-scales for the post-session suggestion. 

```{r}
pcs_items <- c("pcs_handslow", "pcs_mosquito", "pcs_magnetichands", "pcs_armrigidity", "pcs_tastescore", "pcs_postsessionsuggetsion", "pcs_armimmobile", "pcs_music", "pcs_negativehallucination", "pcs_amnesia")

data_ppt$pcs_tastescore <- rowMeans(data_ppt[grepl("taste", names(data_ppt))])
data_ppt$pcs_postsessionsuggetsion <- sqrt(data_ppt$pcs_urgepress * data_ppt$pcs_memorypress)
```



### Illusion Game

```{r}
ig_scores <- data_ig |>
    summarize(
        ErrorRate = sum(Error) / n(),
        IES = mean(RT[Response_Correct == "true"] / mean(Response_Correct == "true")),
        .by = c("Participant", "Illusion_Type", "Condition_Illusion", "Condition_Difficulty")
    ) |>
    pivot_wider(
        names_from = c("Illusion_Type", "Condition_Illusion", "Condition_Difficulty"),
        values_from = c("ErrorRate", "IES")
    )
```

```{r}
corrs <- data.frame()
for (outcome in c("ErrorRate", "IES")) {
    for (ill_type in c("Ebbinghaus", "MullerLyer", "VerticalHorizontal")) {
        for (cond in c("Congruent", "Mild", "Strong")) {
            corrs <- correlation::correlation(
                ig_scores[paste0(outcome, "_", ill_type, "_", cond, "_Easy")],
                ig_scores[paste0(outcome, "_", ill_type, "_", cond, "_Hard")]
            ) |>
                mutate(Outcome = outcome) |>
                rbind(corrs)
        }
    }
}
arrange(corrs, desc(r))

# IF correlations > 0.5:
ig_scores <- data_ig |>
    summarize(
        ErrorRate = sum(Error) / n(),
        IES = mean(RT[Response_Correct == "true"] / mean(Response_Correct == "true")),
        .by = c("Participant", "Illusion_Type", "Condition_Illusion")
    ) |>
    pivot_wider(
        names_from = c("Illusion_Type", "Condition_Illusion"),
        values_from = c("ErrorRate", "IES")
    )
```


#### Outlier Removal

```{r}
#vars <- names(ig_scores)[names(ig_scores) != "Participant"]

#cis <- data.frame()
#for(var in vars){
#  bounds <- bayestestR::hdi(ig_scores[[var]], ci = 0.99)
#  print(paste0(var, ": 95% HDI [", bounds$CI_low, ", ", #bounds$CI_high, "]"))
  
#  ig_scores[[var]] <- ifelse(ig_scores[[var]] > bounds$CI_high, #NA, ig_scores[[var]])
#}

#gt::gt(cis)
```

```{r}
vars <- names(ig_scores)[names(ig_scores) != "Participant"]

ps <- list()
cis <- data.frame()
for(var in vars){
  bounds <- bayestestR::hdi(ig_scores[[var]], ci = 0.95)
  cis <- rbind(cis, data.frame(variable = var, CI_low = bounds$CI_low, CI_high = bounds$CI_high))
  
  # Make plot
  ps[[var]] <- ggplot(ig_scores, aes(x = !!as.name(var))) +
    geom_histogram(bins = 20) +
    geom_vline(xintercept = bounds$CI_low, color = "red") +
    geom_vline(xintercept = bounds$CI_high, color = "red") +
    ggtitle(paste("95% HDI for", var)) 
  
  # Assuming we only want to remove above the top bound
  ig_scores[[var]] <- ifelse(ig_scores[[var]] > bounds$CI_high, NA, ig_scores[[var]])
}

wrap_plots(ps)
gt::gt(cis)
```


#### Manipulation Check

Significant differences between the Congruent and Mild and the Mild and Strong groups are taken as evidence that illusions functioned as intended. 

```{r}
ttests <- data.frame()
for (outcome in c("ErrorRate", "IES")) {
    for (ill_type in c("Ebbinghaus", "MullerLyer", "VerticalHorizontal")) {
        for (cond in c("Congruent", "Mild")) {
            x1 <- ig_scores[[paste0(outcome, "_", ill_type, "_", cond)]]
            if (cond == "Congruent") {
                x2 <- ig_scores[[paste0(outcome, "_", ill_type, "_", "Mild")]]
            } else {
                x2 <- ig_scores[[paste0(outcome, "_", ill_type, "_", "Strong")]]
            }

            complete <- complete.cases(x1, x2)
            x1 <- x1[complete]
            x2 <- x2[complete]

            if (length(x1) < 2) {
                message(paste("Skipping:", outcome, ill_type, cond, "- not enough observations"))
                next
            }

            ttests <- BayesFactor::ttestBF(x = x1, y = x2, paired = TRUE) |>
                parameters::parameters() |>
                mutate(
                    Outcome = outcome,
                    Illusion_Type = ill_type,
                    Condition_Illusion = cond
                ) |>
                rbind(ttests)
        }
    }
}

ttests |>
    select(Illusion_Type, Condition_Illusion, Outcome, Median, BF) |>
    arrange(Illusion_Type, Condition_Illusion, Outcome)
```


## Reliability

### PCS


```{r}
data_ppt[, pcs_items] |>
    psych::alpha()

data_ppt$PCS_Score <- rowMeans(data_ppt[, pcs_items], na.rm = TRUE)
```

### Illusion Game

::: {.panel-tabset}

#### Error Rate - Mild 

```{r}
ig_scores |>
    select(matches("ErrorRate_.*_Mild.*")) |>
    psych::alpha()
```

#### Error Rate - Strong 

```{r}
ig_scores |>
    select(matches("ErrorRate_.*_Strong.*")) |>
    psych::alpha()
```

#### IES - Mild

```{r}
ig_scores |>
    select(matches("IES_.*_Mild.*")) |>
    psych::alpha()
```

#### IES - Strong

```{r}
ig_scores |>
    select(matches("IES_.*_Strong.*")) |>
    psych::alpha()
```

:::


## Illusion Sensitivity vs. PCS 


```{r}
features <- merge(ig_scores, data_ppt[c("Participant", "PCS_Score", "PID5_Psychoticism")])

r_pcs <- correlation::correlation(
    select(features, -PCS_Score, -PID5_Psychoticism),
    select(features, PCS_Score),
    bayesian = TRUE
)

r_pcs |>
    arrange(desc(BF)) |> 
    format_table() |>
    select(-Prior, -Method, -n_Obs, -pd, -`% in ROPE`) |>
    gt::gt() |>
    gt::data_color(
      columns = BF,
      fn = function(bf) dplyr::case_when(
        bf >= 10   ~ "green",
        bf >= 3    ~ "lightgreen",
        bf <= 1/10  ~ "red",
        bf <= 1/3   ~ "#ffabab",
        TRUE       ~ "white"
      )
    )

```

## Illusion Sensitivity vs. Psychoticism 


```{r}
r_psy <- correlation::correlation(
    select(features, -PCS_Score, -PID5_Psychoticism),
    select(features, PID5_Psychoticism),
    bayesian = TRUE
)

r_psy |>
    arrange(desc(BF)) |> 
    format_table() |>
    select(-Prior, -Method, -n_Obs, -pd, -`% in ROPE`) |>
    gt::gt() |>
    gt::data_color(
      columns = BF,
      fn = function(bf) dplyr::case_when(
        bf >= 10   ~ "green",
        bf >= 3    ~ "lightgreen",
        bf <= 1/10  ~ "red",
        bf <= 1/3   ~ "#ffabab",
        TRUE       ~ "white"
      )
    )
```

## Figures

```{r fig1}
#| fig-width: 8
#| fig-height: 6

p <- rbind(mutate(r_pcs, Outcome = "Phenomenological Control"), 
           mutate(r_psy, Outcome = "Psychoticism")) |>
  separate(Parameter1, into = c("Measure", "Illusion", "Condition"), sep = "_") |>
  mutate(Measure = recode(Measure, "ErrorRate" = "Error rate","IES" = "Inverse Efficiency Score (IES)"),
         Illusion = recode(Illusion, "MullerLyer" = "Müller-Lyer","VerticalHorizontal" = "Vertical–Horizontal"),
         Condition = factor(Condition, levels = c("Congruent", "Mild", "Strong"))) |>
  ggplot(aes(x = rho, y = Illusion, color = Condition)) +
  annotate("rect", xmin = -0.1, xmax = 0.1, ymin = -Inf, ymax = Inf, fill = "red", alpha = 0.1) +
  annotate("rect", xmin = 0.1, xmax = Inf, ymin = -Inf, ymax = Inf, fill = "lightgreen", alpha = 0.1) +
  annotate("rect", xmin = -Inf, xmax = -0.1, ymin = -Inf, ymax = Inf, fill = "lightgreen", alpha = 0.1) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  geom_point(position = position_dodge(width = 0.5), size = 3) +
  geom_linerange(aes(xmin = CI_low, xmax = CI_high), position = position_dodge(width = 0.5), linewidth = 1) +
  # geom_label(aes(label = format_bf(BF, stars=TRUE, protect_ratio=F)), 
  #           position = position_dodge(width = 0.5), hjust = 0.5, size = 2, vjust=1) +
  geom_text(aes(x=0, label = format_bf(BF, stars=TRUE, protect_ratio=F)), 
            position = position_dodge(width = 0.5), hjust = 0.5, size = 2, vjust=-0.5, show.legend = FALSE) +
  scale_color_manual(values = c("Congruent" = "#2196F3", "Mild" = "darkorange", "Strong" = "red"),
                     breaks = c("Strong", "Mild", "Congruent")) +
  scale_x_continuous(limits = c(-0.35, 0.35), breaks = c(-0.3, seq(-0.2, 0.2, by = 0.1), 0.3),
                     labels = c("- .30", "- .20", "- .10", "0", ".10", ".20", ".30")) +
  facet_grid(Measure ~ Outcome) +
  labs(x = "Correlation", y = "Illusion", color = "Condition",
       title = "Correlation between Illusion Sensitivity and Dispositional Traits") +
  theme_minimal() +
   theme(axis.text.y = element_text(size = 10, face = "bold"),
         axis.title.y = element_blank(),
         axis.title.x = element_blank(),
         legend.title = element_text(size = 12),
         strip.text.x = element_text(size = 11, face = "bold"),
         strip.text.y = element_text(size = 10, face = "italic"),
         plot.title = element_text(size = 14, face = "bold", hjust = 0.5))
p
```

```{r}
plot(correlation::cor_test(features, "ErrorRate_Ebbinghaus_Strong", "PCS_Score"))
```
