---
title: "Illusion Sensitivity and Phenomenological Control"
subtitle: "Analysis - Departures From Preregistration"
date: "`r Sys.Date()`"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: true
    toc: true
    warning: false 
    message: false 
execute: 
  cache: false
---

## Data Preparation 

```{r}
library(tidyverse)
library(ggdist)
library(ggside)
library(easystats)
library(patchwork)
library(psych)


data_ppt <- read.csv("../data/data_participants.csv")

data_ig <- read.csv("../data/data_illusion.csv") |>
    mutate(
        Block = as.factor(Block),
        Illusion_Effect = fct_relevel(as.factor(Illusion_Effect), "Incongruent", "Congruent"),
        Response_Correct = as.factor(Response_Correct)
    )

```


## Score Computation

### PCS

Mean subjective scale responses for item 4 (Taste) and item 10 (Post-session suggestion) are each generated from the combination of two responses which require processing prior to generating an overall subjective score. 
The subjective score for item 4 is the mean of the responses to the (a)‘sour’ and (b)‘sweet’ sub-scales for the taste suggestion. 
The subjective score for item 10 is the geometric mean (the square root of the product) of the responses to the (a) ‘urge’ and (b) ‘amnesia’ sub-scales for the post-session suggestion. 

```{r}
pcs_items <- c("pcs_handslow", "pcs_mosquito", "pcs_magnetichands", "pcs_armrigidity", "pcs_tastescore", "pcs_postsessionsuggetsion", "pcs_armimmobile", "pcs_music", "pcs_negativehallucination", "pcs_amnesia")

data_ppt$pcs_tastescore <- rowMeans(data_ppt[grepl("taste", names(data_ppt))])
data_ppt$pcs_postsessionsuggetsion <- sqrt(data_ppt$pcs_urgepress * data_ppt$pcs_memorypress)
```



### Illusion Game

```{r}
ig_scores <- data_ig |>
    summarize(
        ErrorRate = sum(Error) / n(),
        IES = mean(RT[Response_Correct == "true"] / mean(Response_Correct == "true")),
        .by = c("Participant", "Illusion_Type", "Condition_Illusion", "Condition_Difficulty")
    ) |>
    pivot_wider(
        names_from = c("Illusion_Type", "Condition_Illusion", "Condition_Difficulty"),
        values_from = c("ErrorRate", "IES")
    )


```

#### Mild vs. Strong

```{r}
corrs <- data.frame()
for (outcome in c("ErrorRate", "IES")) {
    for (ill_type in c("Ebbinghaus", "MullerLyer", "VerticalHorizontal")) {
       for (dif in c("Easy", "Hard")) {
  
         corrs <- correlation::correlation(
                ig_scores[paste0(outcome, "_", ill_type,"_Mild", "_", dif )],
                ig_scores[paste0(outcome, "_", ill_type, "_Strong", "_", dif )]
            ) |>
                mutate(Outcome = outcome) |>
                rbind(corrs)
        }
    }
}
arrange(corrs, desc(r))
```

#### Easy vs. Hard


```{r}
corrs_2 <- data.frame()
for (outcome in c("ErrorRate", "IES")) {
    for (ill_type in c("Ebbinghaus", "MullerLyer", "VerticalHorizontal")) {
        for (cond in c("Congruent", "Mild", "Strong")) {
            corrs_2 <- correlation::correlation(
                ig_scores[paste0(outcome, "_", ill_type, "_", cond, "_Easy")],
                ig_scores[paste0(outcome, "_", ill_type, "_", cond, "_Hard")]
            ) |>
                mutate(Outcome = outcome) |>
                rbind(corrs_2)
        }
    }
}
arrange(corrs_2, desc(r))
```


#### Final Indices

```{r}
ig_scores_incongruent <- data_ig |>
    filter(Condition_Illusion != "Congruent") |>
    summarize(
        ErrorRate = sum(Error) / n(),
        IES = mean(RT[Response_Correct == "true"] / mean(Response_Correct == "true")),
        .by = c("Participant", "Illusion_Type", "Condition_Difficulty")
    ) |>
    pivot_wider(
        names_from = c("Illusion_Type", "Condition_Difficulty"),
        values_from = c("ErrorRate", "IES")
    )
names(ig_scores_incongruent) <- str_replace_all(names(ig_scores_incongruent), "_Hard", "_Incongruent_Hard")
names(ig_scores_incongruent) <- str_replace_all(names(ig_scores_incongruent), "_Easy", "_Incongruent_Easy")
ig_scores <- full_join(ig_scores, ig_scores_incongruent, by = "Participant")


ig_scores_general <- data_ig |>
    summarize(
        ErrorRate = sum(Error) / n(),
        IES = mean(RT[Response_Correct == "true"] / mean(Response_Correct == "true")),
        .by = c("Participant", "Illusion_Type", "Condition_Illusion")
    ) |>
    pivot_wider(
        names_from = c("Illusion_Type", "Condition_Illusion"),
        values_from = c("ErrorRate", "IES")
    ) |> 
  data_addsuffix(select=-Participant, pattern = "_General")
names(ig_scores_general)
ig_scores <- full_join(ig_scores, ig_scores_general, by = "Participant")


ig_scores_incongruent_general <- data_ig |>
    filter(Condition_Illusion != "Congruent") |>
    summarize(
        ErrorRate = sum(Error) / n(),
        IES = mean(RT[Response_Correct == "true"] / mean(Response_Correct == "true")),
        .by = c("Participant", "Illusion_Type")
    ) |>
    pivot_wider(
        names_from = c("Illusion_Type"),
        values_from = c("ErrorRate", "IES")
    ) |> 
   data_addsuffix(select=-Participant, pattern = "_Incongruent_General")

ig_scores <- full_join(ig_scores, ig_scores_incongruent_general, by = "Participant")
```

##### Mild vs strong collapsed across easy and hard

```{r}
corrs_3 <- data.frame()
for (outcome in c("ErrorRate", "IES")) {
    for (ill_type in c("Ebbinghaus", "MullerLyer", "VerticalHorizontal")) {
            corrs_3 <- correlation::correlation(
                ig_scores[paste0(outcome, "_", ill_type, "_Mild_General")],
                ig_scores[paste0(outcome, "_", ill_type, "_Strong_General")]
            ) |>
                mutate(Outcome = outcome) |>
                rbind(corrs_3)
        }
}
arrange(corrs_3, desc(r))
```



##### All General 
```{r}

corrs_4 <- data.frame()
for (outcome in c("ErrorRate", "IES")) {
    for (ill_type in c("Ebbinghaus", "MullerLyer", "VerticalHorizontal")) {
      corrs_4 <- correlation::correlation(
                ig_scores[paste0(outcome, "_", ill_type, "_Congruent_General")],
                ig_scores[paste0(outcome, "_", ill_type, "_Incongruent_General")]
            ) |>
                mutate(Outcome = outcome) |>
                rbind(corrs_4)
    }
}
arrange(corrs_4, desc(r))

```


#### Outlier Removal


```{r}
vars <- names(ig_scores)[names(ig_scores) != "Participant"]
vars <- vars[str_detect(vars, "ErrorRate")]
vars <- vars[str_detect(vars, "General")]
vars <- vars[str_detect(vars, "Congruent|Incongruent")]

outliers <- list("MullerLyer" = c(), "Ebbinghaus" = c(), "VerticalHorizontal" = c())
ps <- list()
cis <- data.frame()
for(var in vars){
    
  bounds <- bayestestR::hdi(ig_scores[[var]], ci = 0.975)
  cis <- rbind(cis, data.frame(variable = var, CI_low = bounds$CI_low, CI_high = bounds$CI_high))

  # Make plot
  ps[[var]] <- ggplot(ig_scores, aes(x = !!as.name(var))) +
    geom_histogram(binwidth = 0.02) +
    geom_vline(xintercept = bounds$CI_low, color = "red") +
    geom_vline(xintercept = bounds$CI_high, color = "red") +
    coord_cartesian(xlim = c(0, 1))  +
    ggtitle(paste("95% HDI for", var))  
  
  illu <- str_split_fixed(var, "_", 4)[,2]
  flagged_ppts <- ig_scores$Participant[(ig_scores[[var]] > bounds$CI_high) | (ig_scores[[var]] < bounds$CI_low) |  is.na(ig_scores[[var]])]
  # print(length(flagged_ppts))
  outliers[[illu]] <- unique(c(outliers[[illu]], flagged_ppts))
}

wrap_plots(ps)
gt::gt(cis)

for(illu in names(outliers)) {
  print(paste0("Outliers for ", illu, ": ", length(outliers[[illu]]), " (",
               format_percent(length(outliers[[illu]]) / nrow(ig_scores)), ")", collapse = ", "))
  ig_scores[ig_scores$Participant %in% outliers[[illu]], str_detect(names(ig_scores), illu)] <- NA
}
```



#### Correlation


```{r}
#| fig-width: 15
#| fig-height: 15

correlation::correlation(select(ig_scores, -Participant), redundant=TRUE) |> 
  cor_sort() |> 
  ggplot(aes(Parameter1, Parameter2, fill = r)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```




#### Manipulation Check

Significant differences between the Congruent and Mild and the Mild and Strong groups are taken as evidence that illusions functioned as intended. 

```{r}
ttests <- data.frame()
for (outcome in c("ErrorRate", "IES")) {
    for (ill_type in c("Ebbinghaus", "MullerLyer", "VerticalHorizontal")) {
        for (cond in c("Congruent", "Mild")) {
            x1 <- ig_scores[[paste0(outcome, "_", ill_type, "_", cond, "_General")]]
            if (cond == "Congruent") {
                x2 <- ig_scores[[paste0(outcome, "_", ill_type, "_", "Mild_General")]]
            } else {
                x2 <- ig_scores[[paste0(outcome, "_", ill_type, "_", "Strong_General")]]
            }

            complete <- complete.cases(x1, x2)
            x1 <- x1[complete]
            x2 <- x2[complete]

            if (length(x1) < 2) {
                message(paste("Skipping:", outcome, ill_type, cond, "- not enough observations"))
                next
            }

            ttests <- BayesFactor::ttestBF(x = x1, y = x2, paired = TRUE) |>
                parameters::parameters() |>
                mutate(
                    Outcome = outcome,
                    Illusion_Type = ill_type,
                    Condition_Illusion = cond
                ) |>
                rbind(ttests)
        }
    }
}

ttests |>
    select(Illusion_Type, Condition_Illusion, Outcome, Median, BF) |>
    arrange(Illusion_Type, Condition_Illusion, Outcome) 


```


## Reliability

### PCS

::: {.panel-tabset}

#### All

```{r}
data_ppt[, pcs_items] |>
    psych::alpha()

data_ppt$PCS_Score <- rowMeans(data_ppt[, pcs_items], na.rm = TRUE)
```

#### PCS3 Excluded

```{r}
data_ppt[data_ppt$PCS_Outlier_Press == F, pcs_items] |>
    psych::alpha()
```


:::


### Illusion Game

::: {.panel-tabset}

#### Error Rate - Mild 

```{r}
ig_scores |>
    select(matches("ErrorRate_.*_Mild.*_General")) |>
    psych::alpha()
```

#### Error Rate - Strong 

```{r}
ig_scores |>
    select(matches("ErrorRate_.*_Strong.*_General")) |>
    psych::alpha()
```

#### IES - Mild

```{r}
ig_scores |>
    select(matches("IES_.*_Mild.*_General")) |>
    psych::alpha()
```

#### IES - Strong

```{r}
ig_scores |>
    select(matches("IES_.*_Strong.*_General")) |>
    psych::alpha()
```

:::


## Correlations

::: {.panel-tabset}

#### All

```{r}

for (illu in names(outliers)) {
  cols <- str_detect(names(ig_scores), illu)
  n_valid <- sum(
    rowSums(!is.na(ig_scores[, cols])) > 0
  )
  print( paste0("Final N for ", illu, ": ", n_valid, " (",format_percent(n_valid / nrow(ig_scores)), ")"))
}
```

The final sample includes `r report::report_participants(data_ppt, age="Age", sex="Sex", education="Education")`, for a total trial number of `r nrow(data_ig)`.


```{r}
make_cor <- function(data) {
  features <- merge(ig_scores, data[c("Participant", "PCS_Score", "PID5_Psychoticism")])
  
  rez <- correlation::correlation(
    dplyr::select(features, -PCS_Score, -PID5_Psychoticism),
    dplyr::select(features, PCS_Score, PID5_Psychoticism),
    bayesian = TRUE
  )
  
  table <- rez |>
    dplyr::arrange(desc(BF)) |> 
    format_table() |>
    dplyr::mutate(BF = as.numeric(BF)) |> 
    dplyr::select(-Prior, -Method, -n_Obs, -pd, -`% in ROPE`) |>
    gt::gt() |>
    gt::data_color(
      columns = BF,
      fn = function(bf) dplyr::case_when(
        bf >= 10   ~ "green",
        bf >= 3    ~ "lightgreen",
        bf <= 1/10 ~ "red",
        bf <= 1/3  ~ "#ffabab",
        .default   = "white"
      )
    )
  
  assign("features", features, envir = .GlobalEnv)
  assign("rez", rez, envir = .GlobalEnv)
  
  table
}

make_cor(data=data_ppt) 

```

#### PCS3 Excluded

```{r}
make_cor(data=data_ppt[data_ppt$PCS_Outlier_Press == F, ])
```


:::

## Figures

```{r fig1}
#| fig-width: 10
#| fig-height: 8

# rez <- rez
# features <- features 

p <- rez |> 
  mutate(Outcome = ifelse(Parameter2 == "PCS_Score", "Phenomenological Control",  "Psychoticism")) |>
  separate(Parameter1, into = c("Measure", "Illusion", "Condition", "Difficulty"), sep = "_") |>
  filter(Difficulty == "General") |>
  mutate(Measure = recode(Measure, "ErrorRate" = "Error rate","IES" = "Inverse Efficiency Score (IES)"),
         Illusion = recode(Illusion, "MullerLyer" = "Müller-Lyer","VerticalHorizontal" = "Vertical–Horizontal"),
         Condition = factor(Condition, levels = c("Congruent", "Mild", "Strong"))) |>
  ggplot(aes(x = rho, y = Illusion, color = Condition)) +
  annotate("rect", xmin = -0.1, xmax = 0.1, ymin = -Inf, ymax = Inf, fill = "red", alpha = 0.1) +
  annotate("rect", xmin = 0.1, xmax = Inf, ymin = -Inf, ymax = Inf, fill = "lightgreen", alpha = 0.1) +
  annotate("rect", xmin = -Inf, xmax = -0.1, ymin = -Inf, ymax = Inf, fill = "lightgreen", alpha = 0.1) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  geom_point(position = position_dodge(width = 0.5), size = 3) +
  geom_linerange(aes(xmin = CI_low, xmax = CI_high), position = position_dodge(width = 0.5), linewidth = 1) +
  # geom_label(aes(label = format_bf(BF, stars=TRUE, protect_ratio=F)), 
  #           position = position_dodge(width = 0.5), hjust = 0.5, size = 2, vjust=1) +
  geom_text(aes(x=0, label = format_bf(BF, stars=TRUE, protect_ratio=F)), 
            position = position_dodge(width = 0.5), hjust = 0.5, size = 2, vjust=-0.5, show.legend = FALSE) +
  scale_color_manual(values = c("Congruent" = "#2196F3", "Mild" = "darkorange", "Strong" = "red"),
                     breaks = c("Strong", "Mild", "Congruent")) +
  scale_x_continuous(limits = c(-0.35, 0.35), breaks = c(-0.3, seq(-0.2, 0.2, by = 0.1), 0.3),
                     labels = c("- .30", "- .20", "- .10", "0", ".10", ".20", ".30")) +
  facet_grid(Measure ~ Outcome) +
  labs(x = "Correlation", y = "Illusion", color = "Condition",
       title = "Correlation between Illusion Sensitivity and Dispositional Traits") +
  theme_minimal() +
   theme(axis.text.y = element_text(size = 10, face = "bold"),
         axis.title.y = element_blank(),
         axis.title.x = element_blank(),
         legend.title = element_text(size = 12),
         strip.text.x = element_text(size = 11, face = "bold"),
         strip.text.y = element_text(size = 10, face = "italic"),
         plot.title = element_text(size = 14, face = "bold", hjust = 0.5))
p
```

## Tables

```{r}
format_bf <- function(bf) {
  ifelse(bf > 1000, "BF₁₀ > 1000",
         sprintf("BF₁₀ = %.2f", bf))
}

ttests_table <- ttests |>
  mutate(
    Condition_Illusion = recode(Condition_Illusion, "Congruent" = "Congruent vs Mild","Mild" = "Mild vs Strong"),
    Illusion_Type = recode( Illusion_Type,"MullerLyer" = "Müller-Lyer","VerticalHorizontal" = "Vertical–Horizontal" )) |>
   select(Illusion_Type, Condition_Illusion, Outcome, Median, BF) |>
  pivot_wider(names_from = c(Outcome, Condition_Illusion),values_from = c(Median, BF)) |>
  arrange(Illusion_Type) |>
  mutate(
    `Error Rate: Congruent vs Mild` = sprintf(
      "M = %.2f, %s",
      `Median_ErrorRate_Congruent vs Mild`,
      format_bf(`BF_ErrorRate_Congruent vs Mild`)),
    `Error Rate: Mild vs Strong` = sprintf(
      "M = %.2f, %s",
     `Median_ErrorRate_Mild vs Strong`,
      format_bf(`BF_ErrorRate_Mild vs Strong`)),
    `IES: Congruent vs Mild` = sprintf(
      "M = %.2f, %s",
      `Median_IES_Congruent vs Mild`,
      format_bf(`BF_IES_Congruent vs Mild`)),
    `IES: Mild vs Strong` = sprintf(
      "M = %.2f, %s",
      `Median_IES_Mild vs Strong`,
      format_bf(`BF_IES_Mild vs Strong`))) |>
  select(Illusion_Type,`Error Rate: Congruent vs Mild`,`Error Rate: Mild vs Strong`,`IES: Congruent vs Mild`,`IES: Mild vs Strong`)


table1 <- ttests_table |>
  gt::gt(rowname_col = "Illusion_Type") |>
  gt::tab_spanner(label = "Error Rate", columns = c(`Error Rate: Congruent vs Mild`, `Error Rate: Mild vs Strong`)) |>
  gt::tab_spanner(label = "IES",columns = c(`IES: Congruent vs Mild`, `IES: Mild vs Strong`) ) |>
  gt::cols_label(Illusion_Type = "Illusion Type") 

gt::gtsave(table1, "table1.tex", expand = 10)

table1
  
```

