---
title: "Illusion Sensitivity and Phenomenological Control"
subtitle: "Analysis - Preregistered"
date: "`r Sys.Date()`"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: true
    toc: true
    warning: false 
    message: false 
execute: 
  cache: false
---

## Data Preparation 

```{r}
library(tidyverse)
library(ggdist)
library(ggside)
library(easystats)
library(patchwork)
library(psych)


data_ppt <- read.csv("../Data/data_participants.csv")

data_ig <- read.csv("../Data/data_illusion.csv") |>
    mutate(
        Block = as.factor(Block),
        Illusion_Effect = fct_relevel(as.factor(Illusion_Effect), "Incongruent", "Congruent"),
        Response_Correct = as.factor(Response_Correct)
    )

```

## PCS Exclusion

```{r}
outliers_pcs1 <- data_ppt[data_ppt$pcs_hello != "Hello", "Participant"]
outliers_pcs2 <- data_ppt[data_ppt$pcs_noballs == TRUE, "Participant"]
outliers_pcs3 <- data_ppt[data_ppt$pcs_press < 5, "Participant"]
outliers_pcs <- unique(c(outliers_pcs1, outliers_pcs2, outliers_pcs3))
```

Using all three Phenomenological Control Scale attention checks, we removed `r length(outliers_pcs)` (`r insight::format_percent(length(outliers_pcs) / nrow(data_ppt))`) participants that either did not report hearing "Hello" at the beginning of the experiment, responded that no balls were presented on the screen, or pressed the space bar fewer than 5 times when asked to do so.


```{r}
data_ppt <- filter(data_ppt, !Participant %in% outliers_pcs)
data_ig <- filter(data_ig, !Participant %in% outliers_pcs)
```

::: {.panel-tabset}

### PCS Exclusions Table

```{r}
pcs_ex_tib <- bind_rows(
  tibble(Participant = outliers_pcs1, Reason = "Failed 'Hello' check"),
  tibble(Participant = outliers_pcs2, Reason = "Reported no balls"),
  tibble(Participant = outliers_pcs3, Reason = "Pressed space bar < 5 times")
)

pcs_ex_tib |> 
  knitr::kable(caption = "Exclusions From Phenomenological Control Scale") |> 
  kableExtra::scroll_box(height = "300px") |> 
  kableExtra::kable_styling()
```

### PCS Exclusions Graph

```{r}
pcs_ex_bar <- pcs_ex_tib |> 
  ggplot(aes(x = Reason, fill = Reason)) +
  geom_bar() +
  scale_x_discrete(labels = scales::label_wrap(17)) +
  scale_y_continuous(breaks = scales::pretty_breaks()) +
    labs(
    title = "Exclusions From Phenomenological Control Scale",
    x = "Reason for Exclusion",
    y = "Count",
  ) +
  theme_minimal() +
  guides(fill = "none")

pcs_ex_bar
```

:::

## Score Computation

### PCS

Mean subjective scale responses for item 4 (Taste) and item 10 (Post-session suggestion) are each generated from the combination of two responses which require processing prior to generating an overall subjective score. 
The subjective score for item 4 is the mean of the responses to the (a)‘sour’ and (b)‘sweet’ sub-scales for the taste suggestion. 
The subjective score for item 10 is the geometric mean (the square root of the product) of the responses to the (a) ‘urge’ and (b) ‘amnesia’ sub-scales for the post-session suggestion. 

```{r}
pcs_items <- c("pcs_handslow", "pcs_mosquito", "pcs_magnetichands", "pcs_armrigidity", "pcs_tastescore", "pcs_postsessionsuggetsion", "pcs_armimmobile", "pcs_music", "pcs_negativehallucination", "pcs_amnesia")

data_ppt$pcs_tastescore <- rowMeans(data_ppt[grepl("taste", names(data_ppt))])
data_ppt$pcs_postsessionsuggetsion <- sqrt(data_ppt$pcs_urgepress * data_ppt$pcs_memorypress)
```



### Illusion Game

::: {.panel-tabset}

#### Outcomes

```{r}
ig_scores <- data_ig |>
    summarize(
        ErrorRate = sum(Error) / n(),
        IES = mean(RT[Response_Correct == "true"] / mean(Response_Correct == "true")),
        .by = c("Participant", "Illusion_Type", "Condition_Illusion", "Condition_Difficulty")
    ) |>
    pivot_wider(
        names_from = c("Illusion_Type", "Condition_Illusion", "Condition_Difficulty"),
        values_from = c("ErrorRate", "IES")
    )
```

```{r}

```

:::

```{r}
corrs <- data.frame()
for (outcome in c("ErrorRate", "IES")) {
    for (ill_type in c("Ebbinghaus", "MullerLyer", "VerticalHorizontal")) {
        for (cond in c("Congruent", "Mild", "Strong")) {
            corrs <- correlation::correlation(
                ig_scores[paste0(outcome, "_", ill_type, "_", cond, "_Easy")],
                ig_scores[paste0(outcome, "_", ill_type, "_", cond, "_Hard")]
            ) |>
                mutate(Outcome = outcome) |>
                rbind(corrs)
        }
    }
}
arrange(corrs, desc(r))

# IF correlations > 0.5:
ig_scores <- data_ig |>
    summarize(
        ErrorRate = sum(Error) / n(),
        IES = mean(RT[Response_Correct == "true"] / mean(Response_Correct == "true")),
        .by = c("Participant", "Illusion_Type", "Condition_Illusion")
    ) |>
    pivot_wider(
        names_from = c("Illusion_Type", "Condition_Illusion"),
        values_from = c("ErrorRate", "IES")
    )
```

```{r}
# collapse across Mild and Strong (Incongruent) conditions

ig_scores_collapsed <- data_ig |>
    mutate(
        Condition_Illusion = if_else(
            Condition_Illusion %in% c("Mild", "Strong"), 
            "Incongruent", 
            Condition_Illusion
        )
    ) |>
    summarize(
        ErrorRate = sum(Error) / n(),
        IES = mean(RT[Response_Correct == "true"] / mean(Response_Correct == "true")),
        .by = c("Participant", "Illusion_Type", "Condition_Illusion")
    ) |>
    pivot_wider(
        names_from = c("Illusion_Type", "Condition_Illusion"),
        values_from = c("ErrorRate", "IES")
    )


corrs_collapsed <- data.frame()
for (outcome in c("ErrorRate", "IES")) {
    for (ill_type in c("Ebbinghaus", "MullerLyer", "VerticalHorizontal")) {
        corrs_collapsed <- correlation::correlation(
            ig_scores_collapsed[paste0(outcome, "_", ill_type, "_Congruent")],
            ig_scores_collapsed[paste0(outcome, "_", ill_type, "_Incongruent")]
        ) |>
            mutate(Outcome = outcome, Illusion_Type = ill_type) |>
            rbind(corrs_collapsed)
    }
}
arrange(corrs_collapsed, desc(r))
```


#### Outlier Removal

::: {.panel-tabset}

#### Outcomes

```{r}
#| fig-width: 20
#| fig-height: 14


vars <- names(ig_scores)[names(ig_scores) != "Participant"]

ps <- list()
cis <- data.frame()
for(var in vars){
  bounds <- bayestestR::hdi(ig_scores[[var]], ci = 0.975)
  cis <- rbind(cis, data.frame(Outcome = var, CI_low = bounds$CI_low, CI_high = bounds$CI_high))
  
  # Make plot
  binwidth <- ifelse(str_detect(var, "ErrorRate"), 0.02, 0.1)
  xlim <- if(str_detect(var, "ErrorRate")) { c(0, 1) } else { c(0, NA) }
  col <- ifelse(str_detect(var, "ErrorRate"), "darkorange", "darkred")
  
n_above <- sum(ig_scores[[var]] > bounds$CI_high, na.rm = TRUE)
pct_above <- format_percent(n_above / nrow(ig_scores))
  
ps[[var]] <- ggplot(ig_scores, aes(x = !!as.name(var))) +
  geom_histogram(binwidth = binwidth, fill = col) +
  geom_vline(xintercept = bounds$CI_low, color = "red") +
  geom_vline(xintercept = bounds$CI_high, color = "red") +
  ggtitle(
    paste0("HDI [", round(bounds$CI_low, 2), ", ", round(bounds$CI_high, 2), "]"),
    subtitle = paste0("Removed: n = ", n_above, " (", pct_above, ")")
  ) +
  coord_cartesian(xlim = xlim) +
  theme(
    plot.title = element_text(size = 8),
    plot.subtitle = element_text(size = 7),
    axis.title = element_text(size = 7),
    axis.text = element_text(size = 6)
  )
  
  # Assuming we only want to remove above the top bound
  ig_scores[[var]] <- ifelse(ig_scores[[var]] > bounds$CI_high, NA, ig_scores[[var]])
}

wrap_plots(ps)
gt::gt(cis)
```

#### Outcomes Collapsed

```{r}
#| fig-width: 20
#| fig-height: 14


vars <- names(ig_scores_collapsed)[names(ig_scores_collapsed) != "Participant"]

ps <- list()
cis <- data.frame()
for(var in vars){
  bounds <- bayestestR::hdi(ig_scores_collapsed[[var]], ci = 0.975)
  cis <- rbind(cis, data.frame(Outcome = var, CI_low = bounds$CI_low, CI_high = bounds$CI_high))
  
  # Make plot
  binwidth <- ifelse(str_detect(var, "ErrorRate"), 0.02, 0.1)
  xlim <- if(str_detect(var, "ErrorRate")) { c(0, 1) } else { c(0, NA) }
  col <- ifelse(str_detect(var, "ErrorRate"), "darkorange", "darkred")
  
n_above <- sum(ig_scores_collapsed[[var]] > bounds$CI_high, na.rm = TRUE)
pct_above <- format_percent(n_above / nrow(ig_scores_collapsed))
  
ps[[var]] <- ggplot(ig_scores_collapsed, aes(x = !!as.name(var))) +
  geom_histogram(binwidth = binwidth, fill = col) +
  geom_vline(xintercept = bounds$CI_low, color = "red") +
  geom_vline(xintercept = bounds$CI_high, color = "red") +
  ggtitle(
    paste0("HDI [", round(bounds$CI_low, 2), ", ", round(bounds$CI_high, 2), "]"),
    subtitle = paste0("Removed: n = ", n_above, " (", pct_above, ")")
  ) +
  coord_cartesian(xlim = xlim) +
  theme(
    plot.title = element_text(size = 8),
    plot.subtitle = element_text(size = 7),
    axis.title = element_text(size = 7),
    axis.text = element_text(size = 6)
  )
  
  # Assuming we only want to remove above the top bound
  ig_scores_collapsed[[var]] <- ifelse(ig_scores_collapsed[[var]] > bounds$CI_high, NA, ig_scores_collapsed[[var]])
}

wrap_plots(ps)
gt::gt(cis)
```

:::

#### Manipulation Check

Significant differences between the Congruent and Mild and the Mild and Strong groups are taken as evidence that illusions functioned as intended.

::: {.panel-tabset}

#### Outcomes

```{r}
ttests <- data.frame()
for (outcome in c("ErrorRate", "IES")) {
    for (ill_type in c("Ebbinghaus", "MullerLyer", "VerticalHorizontal")) {
        for (cond in c("Congruent", "Mild")) {
          
            # Add loop for Hard and Easy if need be

            x1 <- ig_scores[[paste0(outcome, "_", ill_type, "_", cond)]]
            if (cond == "Congruent") {
                x2 <- ig_scores[[paste0(outcome, "_", ill_type, "_", "Mild")]]
            } else {
                x2 <- ig_scores[[paste0(outcome, "_", ill_type, "_", "Strong")]]
            }

            complete <- complete.cases(x1, x2)
            x1 <- x1[complete]
            x2 <- x2[complete]

            if (length(x1) < 2) {
                message(paste("Skipping:", outcome, ill_type, cond, "- not enough observations"))
                next
            }

            ttests <- BayesFactor::ttestBF(x = x1, y = x2, paired = TRUE) |>
                parameters::parameters() |>
                mutate(
                    Outcome = outcome,
                    Illusion_Type = ill_type,
                    Condition_Illusion = cond
                ) |>
                rbind(ttests)
        }
    }
}

ttests |>
    select(Illusion_Type, Condition_Illusion, Outcome, Median, BF) |>
    arrange(Illusion_Type, Condition_Illusion, Outcome)
```

#### Outcomes Collapsed

```{r}
ttests_collapsed <- data.frame()
for (outcome in c("ErrorRate", "IES")) {
    for (ill_type in c("Ebbinghaus", "MullerLyer", "VerticalHorizontal")) {

        x1 <- ig_scores_collapsed[[paste0(outcome, "_", ill_type, "_Congruent")]]
        x2 <- ig_scores_collapsed[[paste0(outcome, "_", ill_type, "_Incongruent")]]

        complete <- complete.cases(x1, x2)
        x1 <- x1[complete]
        x2 <- x2[complete]

        if (length(x1) < 2) {
            message(paste("Skipping:", outcome, ill_type, "- not enough observations"))
            next
        }

        ttests_collapsed <- BayesFactor::ttestBF(x = x1, y = x2, paired = TRUE) |>
            parameters::parameters() |>
            mutate(
                Outcome = outcome,
                Illusion_Type = ill_type
            ) |>
            rbind(ttests_collapsed)
    }
}

ttests_collapsed |>
    select(Illusion_Type, Outcome, Median, BF) |>
    arrange(Illusion_Type, Outcome)
```

:::

## Reliability

### PCS


```{r}
data_ppt[, pcs_items] |>
    psych::alpha()

data_ppt$PCS_Score <- rowMeans(data_ppt[, pcs_items], na.rm = TRUE)
```

### Illusion Game

::: {.panel-tabset}

#### Error Rate - Mild 

```{r}
ig_scores |>
    select(matches("ErrorRate_.*_Mild.*")) |>
    psych::alpha()
```

#### Error Rate - Strong 

```{r}
ig_scores |>
    select(matches("ErrorRate_.*_Strong.*")) |>
    psych::alpha()
```

#### IES - Mild

```{r}
ig_scores |>
    select(matches("IES_.*_Mild.*")) |>
    psych::alpha()
```

#### IES - Strong

```{r}
ig_scores |>
    select(matches("IES_.*_Strong.*")) |>
    psych::alpha()
```

:::


## Illusion Sensitivity vs. PCS 


```{r}
features <- merge(ig_scores, data_ppt[c("Participant", "PCS_Score", "PID5_Psychoticism")])

r_pcs <- correlation::correlation(
    select(features, -PCS_Score, -PID5_Psychoticism),
    select(features, PCS_Score),
    bayesian = TRUE
)

r_pcs |>
    arrange(desc(BF)) |> 
    format_table() |>
    select(-Prior, -Method, -n_Obs, -pd, -`% in ROPE`) |>
    gt::gt() |>
    gt::data_color(
      columns = BF,
      fn = function(bf) dplyr::case_when(
        bf >= 10   ~ "green",
        bf >= 3    ~ "lightgreen",
        bf <= 1/10  ~ "red",
        bf <= 1/3   ~ "#ffabab",
        TRUE       ~ "white"
      )
    )

```

## Illusion Sensitivity vs. Psychoticism 


```{r}
r_psy <- correlation::correlation(
    select(features, -PCS_Score, -PID5_Psychoticism),
    select(features, PID5_Psychoticism),
    bayesian = TRUE
)

r_psy |>
    arrange(desc(BF)) |> 
    format_table() |>
    select(-Prior, -Method, -n_Obs, -pd, -`% in ROPE`) |>
    gt::gt() |>
    gt::data_color(
      columns = BF,
      fn = function(bf) dplyr::case_when(
        bf >= 10   ~ "green",
        bf >= 3    ~ "lightgreen",
        bf <= 1/10  ~ "red",
        bf <= 1/3   ~ "#ffabab",
        TRUE       ~ "white"
      )
    )
```

## Figures

```{r fig1}
#| fig-width: 8
#| fig-height: 6

p <- rbind(mutate(r_pcs, Outcome = "Phenomenological Control"), 
           mutate(r_psy, Outcome = "Psychoticism")) |>
  separate(Parameter1, into = c("Measure", "Illusion", "Condition"), sep = "_") |>
  mutate(Measure = recode(Measure, "ErrorRate" = "Error rate","IES" = "Inverse Efficiency Score (IES)"),
         Illusion = recode(Illusion, "MullerLyer" = "Müller-Lyer","VerticalHorizontal" = "Vertical–Horizontal"),
         Condition = factor(Condition, levels = c("Congruent", "Mild", "Strong"))) |>
  ggplot(aes(x = rho, y = Illusion, color = Condition)) +
  annotate("rect", xmin = -0.1, xmax = 0.1, ymin = -Inf, ymax = Inf, fill = "red", alpha = 0.1) +
  annotate("rect", xmin = 0.1, xmax = Inf, ymin = -Inf, ymax = Inf, fill = "lightgreen", alpha = 0.1) +
  annotate("rect", xmin = -Inf, xmax = -0.1, ymin = -Inf, ymax = Inf, fill = "lightgreen", alpha = 0.1) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  geom_point(position = position_dodge(width = 0.5), size = 3) +
  geom_linerange(aes(xmin = CI_low, xmax = CI_high), position = position_dodge(width = 0.5), linewidth = 1) +
  # geom_label(aes(label = format_bf(BF, stars=TRUE, protect_ratio=F)), 
  #           position = position_dodge(width = 0.5), hjust = 0.5, size = 2, vjust=1) +
  geom_text(aes(x=0, label = format_bf(BF, stars=TRUE, protect_ratio=F)), 
            position = position_dodge(width = 0.5), hjust = 0.5, size = 2, vjust=-0.5, show.legend = FALSE) +
  scale_color_manual(values = c("Congruent" = "#2196F3", "Mild" = "darkorange", "Strong" = "red"),
                     breaks = c("Strong", "Mild", "Congruent")) +
  scale_x_continuous(limits = c(-0.35, 0.35), breaks = c(-0.3, seq(-0.2, 0.2, by = 0.1), 0.3),
                     labels = c("- .30", "- .20", "- .10", "0", ".10", ".20", ".30")) +
  facet_grid(Measure ~ Outcome) +
  labs(x = "Correlation", y = "Illusion", color = "Condition",
       title = "Correlation between Illusion Sensitivity and Dispositional Traits") +
  theme_minimal() +
   theme(axis.text.y = element_text(size = 10, face = "bold"),
         axis.title.y = element_blank(),
         axis.title.x = element_blank(),
         legend.title = element_text(size = 12),
         strip.text.x = element_text(size = 11, face = "bold"),
         strip.text.y = element_text(size = 10, face = "italic"),
         plot.title = element_text(size = 14, face = "bold", hjust = 0.5))
p
```

