---
title: "Illusion Sensitivity and Phenomenological Control"
subtitle: "Analysis - Departures From Preregistration"
date: "`r Sys.Date()`"
editor: source
editor_options: 
  chunk_output_type: console
format:
  html:
    code-fold: true
    self-contained: true
    toc: true
    warning: false 
    message: false 
execute: 
  cache: false
---

## Data Preparation 

```{r}
library(tidyverse)
library(ggdist)
library(ggside)
library(easystats)
library(patchwork)
library(psych)


data_ppt <- read.csv("../data/data_participants.csv")

data_ig <- read.csv("../data/data_illusion.csv") |>
    mutate(
        Block = as.factor(Block),
        Illusion_Effect = fct_relevel(as.factor(Illusion_Effect), "Incongruent", "Congruent"),
        Response_Correct = as.factor(Response_Correct)
    )

```



## PCS Score

### Computation

Mean subjective scale responses for item 4 (Taste) and item 10 (Post-session suggestion) are each generated from the combination of two responses which require processing prior to generating an overall subjective score. 
The subjective score for item 4 is the mean of the responses to the (a)‘sour’ and (b)‘sweet’ sub-scales for the taste suggestion. 
The subjective score for item 10 is the geometric mean (the square root of the product) of the responses to the (a) ‘urge’ and (b) ‘amnesia’ sub-scales for the post-session suggestion. 

```{r}
pcs_items <- c("pcs_handslow", "pcs_mosquito", "pcs_magnetichands", "pcs_armrigidity", "pcs_tastescore", "pcs_postsessionsuggetsion", "pcs_armimmobile", "pcs_music", "pcs_negativehallucination", "pcs_amnesia")

data_ppt$pcs_tastescore <- rowMeans(data_ppt[grepl("taste", names(data_ppt))])
data_ppt$pcs_postsessionsuggetsion <- sqrt(data_ppt$pcs_urgepress * data_ppt$pcs_memorypress)
```



## Illusion Game Scores

### Computation 

#### Full Stratification

```{r}
ig_scores <- data_ig |>
    summarize(
        ErrorRate = sum(Error) / n(),
        IES = mean(RT[Response_Correct == "true"] / mean(Response_Correct == "true")),
        .by = c("Participant", "Illusion_Type", "Condition_Illusion", "Condition_Difficulty")
    ) |>
    pivot_wider(
        names_from = c("Illusion_Type", "Condition_Illusion", "Condition_Difficulty"),
        values_from = c("ErrorRate", "IES")
    )
```

#### Collapse Mild and Strong


```{r}
ig_scores_incongruent <- data_ig |>
    filter(Condition_Illusion != "Congruent") |>
    summarize(
        ErrorRate = sum(Error) / n(),
        IES = mean(RT[Response_Correct == "true"] / mean(Response_Correct == "true")),
        .by = c("Participant", "Illusion_Type", "Condition_Difficulty")
    ) |>
    pivot_wider(
        names_from = c("Illusion_Type", "Condition_Difficulty"),
        values_from = c("ErrorRate", "IES")
    )
names(ig_scores_incongruent) <- str_replace_all(names(ig_scores_incongruent), "_Hard", "_Incongruent_Hard")
names(ig_scores_incongruent) <- str_replace_all(names(ig_scores_incongruent), "_Easy", "_Incongruent_Easy")
ig_scores <- full_join(ig_scores, ig_scores_incongruent, by = "Participant")
```


#### Collapse Easy and Hard


```{r}
ig_scores_general <- data_ig |>
    summarize(
        ErrorRate = sum(Error) / n(),
        IES = mean(RT[Response_Correct == "true"] / mean(Response_Correct == "true")),
        .by = c("Participant", "Illusion_Type", "Condition_Illusion")
    ) |>
    pivot_wider(
        names_from = c("Illusion_Type", "Condition_Illusion"),
        values_from = c("ErrorRate", "IES")
    ) |> 
  data_addsuffix(select=-Participant, pattern = "_General")

ig_scores <- full_join(ig_scores, ig_scores_general, by = "Participant")


ig_scores_incongruent_general <- data_ig |>
    filter(Condition_Illusion != "Congruent") |>
    summarize(
        ErrorRate = sum(Error) / n(),
        IES = mean(RT[Response_Correct == "true"] / mean(Response_Correct == "true")),
        .by = c("Participant", "Illusion_Type")
    ) |>
    pivot_wider(
        names_from = c("Illusion_Type"),
        values_from = c("ErrorRate", "IES")
    ) |> 
   data_addsuffix(select=-Participant, pattern = "_Incongruent_General")

ig_scores <- full_join(ig_scores, ig_scores_incongruent_general, by = "Participant")
```



### Outlier Removal

#### Exclusion

```{r}
#| fig-width: 15
#| fig-height: 10

vars <- names(ig_scores)[names(ig_scores) != "Participant"]
vars <- vars[str_detect(vars, "ErrorRate")]
vars <- vars[str_detect(vars, "General")]
vars <- vars[str_detect(vars, "Congruent|Incongruent")]

outliers <- list("MullerLyer" = c(), "Ebbinghaus" = c(), "VerticalHorizontal" = c())
ps <- list()
cis <- data.frame()
for(var in vars){
    
  bounds <- bayestestR::hdi(ig_scores[[var]], ci = 0.975)
  cis <- rbind(cis, data.frame(variable = var, CI_low = bounds$CI_low, CI_high = bounds$CI_high))

  # Make plot
  ps[[var]] <- ggplot(ig_scores, aes(x = !!as.name(var))) +
    geom_histogram(binwidth = 0.015) +
    geom_vline(xintercept = bounds$CI_low, color = "red") +
    geom_vline(xintercept = bounds$CI_high, color = "red") +
    coord_cartesian(xlim = c(0, 1))  +
    ggtitle(paste("97.5% HDI for", var))   +
    theme_minimal()
  
  illu <- str_split_fixed(var, "_", 4)[,2]
  flagged_ppts <- ig_scores$Participant[(ig_scores[[var]] > bounds$CI_high) | (ig_scores[[var]] < bounds$CI_low) |  is.na(ig_scores[[var]])]
  # print(length(flagged_ppts))
  outliers[[illu]] <- unique(c(outliers[[illu]], flagged_ppts))
}

wrap_plots(ps)
gt::gt(cis)

for(illu in names(outliers)) {
  print(paste0("Outliers for ", illu, ": ", length(outliers[[illu]]), " (",
               format_percent(length(outliers[[illu]]) / nrow(ig_scores)), ")", collapse = ", "))
  ig_scores[ig_scores$Participant %in% outliers[[illu]], str_detect(names(ig_scores), illu)] <- NA
  print(paste0("N = ", nrow(ig_scores[!ig_scores$Participant %in% outliers[[illu]], str_detect(names(ig_scores), illu)])))
}
```

#### Correlation


```{r}
#| fig-width: 15
#| fig-height: 15

correlation::correlation(select(ig_scores, -Participant), redundant=TRUE) |> 
  cor_sort() |> 
  ggplot(aes(Parameter1, Parameter2, fill = r)) +
  geom_tile() +
  scale_fill_gradient2(low = "blue", mid = "white", high = "red", midpoint = 0) +
  theme_minimal() +
  theme(axis.text.x = element_text(angle = 45, hjust = 1))


```



### Should we collapse across Mild and Strong, and Easy and Hard?

::: {.panel-tabset}

#### Mild vs. Strong


```{r}
corrs <- data.frame()
for (outcome in c("ErrorRate", "IES")) {
    for (ill_type in c("Ebbinghaus", "MullerLyer", "VerticalHorizontal")) {
       for (dif in c("Easy", "Hard")) {
  
         corrs <- correlation::correlation(
                ig_scores[paste0(outcome, "_", ill_type,"_Mild", "_", dif )],
                ig_scores[paste0(outcome, "_", ill_type, "_Strong", "_", dif )]
            ) |>
                mutate(Illusion = ill_type,
                       Outcome = outcome) |>
                rbind(corrs)
        }
    }
}

corrs2 <- data.frame()
for (outcome in c("ErrorRate", "IES")) {
    for (ill_type in c("Ebbinghaus", "MullerLyer", "VerticalHorizontal")) {
            corrs2 <- correlation::correlation(
                ig_scores[paste0(outcome, "_", ill_type, "_Mild_General")],
                ig_scores[paste0(outcome, "_", ill_type, "_Strong_General")]
            ) |>
                mutate(Illusion = ill_type,
                       Outcome = outcome) |>
                rbind(corrs2)
        }
}



print_corrs <- function(t) {
  arrange(t, Illusion, Outcome, desc(r)) |> 
    mutate(Parameter1 = str_replace(Parameter1, "ErrorRate_|IES_", ""),
           Parameter1 = str_replace(Parameter1, paste0(Illusion, "_"), ""),
           Parameter2 = str_replace(Parameter2, "ErrorRate_|IES_", ""),
           Parameter2 = str_replace(Parameter2, paste0(Illusion, "_"), "")) |> 
    datawizard::data_reorder(c("Illusion", "Outcome")) |> 
    format_table() |> 
    mutate(r = as.numeric(r)) |> 
    gt::gt() |>
    gt::data_color(
        columns = r,
        fn = function(x) dplyr::case_when(
          x >= 0.5  ~ "lightgreen",
          .default   = "#ffabab"
        )
      )
}

print_corrs(rbind(corrs, corrs2))
```

#### Easy vs. Hard

```{r}
corrs3 <- data.frame()
for (outcome in c("ErrorRate", "IES")) {
    for (ill_type in c("Ebbinghaus", "MullerLyer", "VerticalHorizontal")) {
        for (cond in c("Congruent", "Mild", "Strong")) {
            corrs3 <- correlation::correlation(
                ig_scores[paste0(outcome, "_", ill_type, "_", cond, "_Easy")],
                ig_scores[paste0(outcome, "_", ill_type, "_", cond, "_Hard")]
            ) |>
                mutate(Illusion = ill_type,
                       Outcome = outcome) |>
                rbind(corrs3)
        }
    }
}

print_corrs(rbind(corrs3))
```



:::




#### Manipulation Check

Significant differences between the Congruent and Mild and the Mild and Strong groups are taken as evidence that illusions functioned as intended. 

```{r}
ttests <- data.frame()
for (outcome in c("ErrorRate", "IES")) {
    for (ill_type in c("Ebbinghaus", "MullerLyer", "VerticalHorizontal")) {
        for (cond in c("Congruent", "Mild")) {
            x1 <- ig_scores[[paste0(outcome, "_", ill_type, "_", cond, "_General")]]
            if (cond == "Congruent") {
                x2 <- ig_scores[[paste0(outcome, "_", ill_type, "_", "Mild_General")]]
            } else {
                x2 <- ig_scores[[paste0(outcome, "_", ill_type, "_", "Strong_General")]]
            }

            complete <- complete.cases(x1, x2)
            x1 <- x1[complete]
            x2 <- x2[complete]

            if (length(x1) < 2) {
                message(paste("Skipping:", outcome, ill_type, cond, "- not enough observations"))
                next
            }

            ttests <- BayesFactor::ttestBF(x = x1, y = x2, paired = TRUE) |>
                parameters::parameters() |>
                mutate(
                    Outcome = outcome,
                    Illusion_Type = ill_type,
                    Condition_Illusion = cond
                ) |>
                rbind(ttests)
        }
    }
}

ttests |>
    select(Illusion_Type, Condition_Illusion, Outcome, Median, BF) |>
    arrange(Illusion_Type, Condition_Illusion, Outcome) |> 
    format_table() |> 
    gt::gt()
```



```{r}
#| code-fold: true

gt_apastyle <- function(gt_table, font.size=12) {
  gt_table  |> 
    gt::opt_table_lines(extent = "none") %>%
    gt::tab_options(
      heading.border.bottom.width = 2,
      heading.border.bottom.color = "black",
      heading.border.bottom.style = "solid",
      table.border.top.color = "black",
      table.border.top.style = "solid",
      table.border.top.width = 2,  
      table_body.hlines.color = "white",
      table_body.border.top.color = "black",
      table_body.border.top.style = "solid",
      table_body.border.top.width = 2,
      heading.title.font.size = font.size,
      table.font.size = font.size,
      heading.subtitle.font.size = font.size,
      table_body.border.bottom.color = "black",
      table_body.border.bottom.width = 2,
      table_body.border.bottom.style = "solid",
      column_labels.border.bottom.color = "black",
      column_labels.border.bottom.style = "solid",
      column_labels.border.bottom.width = 1,
      latex.use_longtable = FALSE
    ) |> 
      gt::opt_table_font(font = "times")
}


ig_means <- ig_scores |>
  select(contains("Mild_General"), contains("Strong_General"), contains("_Congruent_General")) |>
  describe_distribution()

ig_means_table <- ig_means |>
  separate(Variable,into = c("Outcome", "Illusion_Type", "Condition", "Extra"),sep = "_") |>
  select(Outcome, Illusion_Type, Condition, Mean, SD, Min, Max)|>
  pivot_wider(
    names_from = c(Outcome,  Condition),
    values_from = c(Mean, SD, Min, Max)) |>
  arrange(Illusion_Type) |>
  mutate(`Error Rate Mild` = paste0(insight::format_percent(Mean_ErrorRate_Mild), " ± ",  round(SD_ErrorRate_Mild*100, 0), "%"),
         `Error Rate Strong` = paste0(insight::format_percent(Mean_ErrorRate_Strong)," ± ",  round(SD_ErrorRate_Strong*100, 0), "%"),
         `Error Rate Congruent` = paste0(insight::format_percent(Mean_ErrorRate_Congruent)," ± ",  round(SD_ErrorRate_Congruent*100, 0), "%"),
        
         `IES Mild` = sprintf("%.2f  ± %.2f", Mean_IES_Mild,SD_IES_Mild),
         `IES Strong` = sprintf("%.2f  ± %.2f", Mean_IES_Strong,SD_IES_Strong),
         `IES Congruent` = sprintf("%.2f  ± %.2f", Mean_IES_Congruent,SD_IES_Congruent)) |>
 select(Illusion_Type,`Error Rate Congruent`, `Error Rate Mild`,`Error Rate Strong`, `IES Congruent`, `IES Mild`,`IES Strong` ) |>
  mutate(Illusion_Type = recode(Illusion_Type,"MullerLyer" = "Müller-Lyer","VerticalHorizontal" = "Vertical–Horizontal" ))


table1 <- ig_means_table |>
 gt::gt(rowname_col = "Illusion_Type") |>
  gt::tab_spanner(label = "Error Rate",columns = starts_with("Error Rate")) |>
  gt::tab_spanner(label = "IES",columns = starts_with("IES")) |>
  gt::cols_label(
    `Error Rate Congruent` = "Congruent",
    `Error Rate Mild` = "Mild",
    `Error Rate Strong` = "Strong",
    `IES Congruent` = "Congruent",
    `IES Mild` = "Mild",
    `IES Strong` = "Strong") |>
  gt::tab_source_note(
    source_note = "Note. Values represent means and standard deviations. All Bayes factors (BFs) for the group differences exceeded 1000, indicating extreme evidence in favour of a difference between groups."
  ) |> gt_apastyle()


table1

saveRDS(table1, "figures/table1.rds")
   
```




## Reliability

### PCS

::: {.panel-tabset}

#### All

```{r}
data_ppt[, pcs_items] |>
    psych::alpha()

data_ppt$PCS_Score <- rowMeans(data_ppt[, pcs_items], na.rm = TRUE)
```

#### PCS3 Excluded

```{r}
data_ppt[data_ppt$PCS_Outlier_Press == F, pcs_items] |>
    psych::alpha()
```


:::


### Illusion Game

::: {.panel-tabset}

#### Error Rate - Mild 

```{r}
ig_scores |>
    select(matches("ErrorRate_.*_Mild.*_General")) |>
    psych::alpha()
```

#### Error Rate - Strong 

```{r}
ig_scores |>
    select(matches("ErrorRate_.*_Strong.*_General")) |>
    psych::alpha()
```

#### IES - Mild

```{r}
ig_scores |>
    select(matches("IES_.*_Mild.*_General")) |>
    psych::alpha()
```

#### IES - Strong

```{r}
ig_scores |>
    select(matches("IES_.*_Strong.*_General")) |>
    psych::alpha()
```

:::


## Correlations

::: {.panel-tabset}

#### All

The final sample includes `r report::report_participants(data_ppt, age="Age", sex="Sex", education="Education")`, for a total trial number of `r nrow(data_ig)`.


```{r}
make_cor <- function(data) {
  features <- merge(ig_scores, data[c("Participant", "PCS_Score", "PID5_Psychoticism")])
  
  rez <- correlation::correlation(
    dplyr::select(features, -PCS_Score, -PID5_Psychoticism),
    dplyr::select(features, PCS_Score, PID5_Psychoticism),
    bayesian = TRUE
  )
  
  table <- rez |>
    dplyr::arrange(desc(BF)) |> 
    format_table() |>
    dplyr::mutate(BF = as.numeric(BF),
                  Parameter2 = str_remove_all(Parameter2, "_Score|PID5_")) |> 
    dplyr::select(-Prior, -Method, -n_Obs, -`% in ROPE`) |>
    gt::gt() |>
    gt::data_color(
      columns = BF,
      fn = function(bf) dplyr::case_when(
        bf >= 10   ~ "green",
        bf >= 3    ~ "lightgreen",
        bf <= 1/10 ~ "red",
        bf <= 1/3  ~ "#ffabab",
        .default   = "white"
      )
    )
  
  assign("features", features, envir = .GlobalEnv)
  assign("rez", rez, envir = .GlobalEnv)
  
  table
}

make_cor(data=data_ppt) 

```

#### PCS3 Excluded

```{r}
make_cor(data=data_ppt[data_ppt$PCS_Outlier_Press == F, ])
```


:::

## Figures

```{r fig1}
#| fig-width: 10
#| fig-height: 8
#| dpi: 300

# rez <- rez
# features <- features 

p <- rez |> 
  mutate(Outcome = ifelse(Parameter2 == "PCS_Score", "Phenomenological Control",  "Psychoticism")) |>
  separate(Parameter1, into = c("Measure", "Illusion", "Condition", "Difficulty"), sep = "_") |>
  filter(Difficulty == "General") |>
  mutate(Measure = recode(Measure, "ErrorRate" = "Error rate","IES" = "Inverse Efficiency Score (IES)"),
         Illusion = recode(Illusion, "MullerLyer" = "Müller-Lyer","VerticalHorizontal" = "Vertical–Horizontal"),
         Condition = factor(Condition, levels = c("Congruent", "Mild", "Strong"))) |>
  ggplot(aes(x = rho, y = Illusion, color = Condition)) +
  annotate("rect", xmin = -0.1, xmax = 0.1, ymin = -Inf, ymax = Inf, fill = "red", alpha = 0.1) +
  annotate("rect", xmin = 0.1, xmax = Inf, ymin = -Inf, ymax = Inf, fill = "lightgreen", alpha = 0.1) +
  annotate("rect", xmin = -Inf, xmax = -0.1, ymin = -Inf, ymax = Inf, fill = "lightgreen", alpha = 0.1) +
  geom_vline(xintercept = 0, linetype = "dashed") +
  geom_point(position = position_dodge(width = 0.5), size = 3) +
  geom_linerange(aes(xmin = CI_low, xmax = CI_high), position = position_dodge(width = 0.5), linewidth = 1) +
  # geom_label(aes(label = format_bf(BF, stars=TRUE, protect_ratio=F)), 
  #           position = position_dodge(width = 0.5), hjust = 0.5, size = 2, vjust=1) +
  geom_text(aes(x=0, label = format_bf(BF, stars=TRUE, protect_ratio=F)), 
            position = position_dodge(width = 0.5), hjust = 0.5, size = 2, vjust=-0.5, show.legend = FALSE) +
  scale_color_manual(values = c("Congruent" = "#2196F3", "Mild" = "darkorange", "Strong" = "red"),
                     breaks = c("Strong", "Mild", "Congruent")) +
  scale_x_continuous(limits = c(-0.35, 0.35), breaks = c(-0.3, seq(-0.2, 0.2, by = 0.1), 0.3),
                     labels = c("- .30", "- .20", "- .10", "0", ".10", ".20", ".30")) +
  facet_grid(Measure ~ Outcome) +
  labs(x = "Correlation", y = "Illusion", color = "Condition",
       title = "Correlation between Illusion Sensitivity and Dispositional Traits") +
  theme_minimal() +
   theme(axis.text.y = element_text(size = 10, face = "bold"),
         axis.title.y = element_blank(),
         axis.title.x = element_blank(),
         legend.title = element_text(size = 12),
         strip.text.x = element_text(size = 11, face = "bold"),
         strip.text.y = element_text(size = 10, face = "italic"),
         plot.title = element_text(size = 14, face = "bold", hjust = 0.5))
p
```


